---
title: 'Supervised Learning : Exploring Boosting, Bagging and Stacking Algorithms for Ensemble Learning'
author: "John Pauline Pineda"
date: "August 19, 2023"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    highlight: tango
    css: doc.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=15, fig.height=10)
```

# **1. Table of Contents**
|
| Details.
|
| Details.
|
##  1.1 Sample Data
|
|
| Preliminary dataset assessment:
|
| 

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.1, warning=FALSE, message=FALSE}

```

</details>

##  1.2 Data Quality Assessment
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.2, warning=FALSE, message=FALSE}

```

</details>

##  1.3 Data Preprocessing

###  1.3.1 Outlier Detection
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.1.1, warning=FALSE, message=FALSE}

```

```{r section_1.3.1.2, warning=FALSE, message=FALSE, fig.width=15, fig.height=2}

```

</details>

###  1.3.2 Zero and Near-Zero Variance
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.2, warning=FALSE, message=FALSE}

```

</details>

###  1.3.3 Collinearity
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.3, warning=FALSE, message=FALSE}

```

</details>

###  1.3.4 Linear Dependency
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.4, warning=FALSE, message=FALSE}

```

</details>

###  1.3.5 Distributional Shape
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.3.5, warning=FALSE, message=FALSE, fig.width=15, fig.height=2}

```

</details>

## 1.4 Data Exploration
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.4, warning=FALSE, message=FALSE}

```

</details>

## 1.5 Model Boosting

###  1.5.1 Boosting Model 1
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.1, warning=FALSE, message=FALSE}

```

</details>

###  1.5.2 Boosting Model 2
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.2, warning=FALSE, message=FALSE}

```

</details>

###  1.5.3 Boosting Model 3
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.5.3, warning=FALSE, message=FALSE}

```

</details>

## 1.6 Model Bagging

###  1.6.1 Bagging Model 1
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.6.1, warning=FALSE, message=FALSE}

```

</details>

###  1.6.2 Boosting Model 2
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.6.2, warning=FALSE, message=FALSE}

```

</details>

## 1.7 Model Stacking

###  1.7.1 Stacking Model 1
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.7.1, warning=FALSE, message=FALSE}

```

</details>

##  1.8 Algorithm Comparison Summary
|
| Details.
|
|

<details><summary><mark style="background-color: #000000;color: #FFFFFF">**Code Chunk | Output**</mark></summary>

```{r section_1.8, warning=FALSE, message=FALSE, dev='png'}

```

</details>

|
# **2. References**
|
| **[Book]** [Applied Predictive Modeling](http://appliedpredictivemodeling.com/) by Max Kuhn and Kjell Johnson
| **[Book]** [An Introduction to Statistical Learning](https://www.statlearning.com/) by Gareth James, Daniela Witten, Trevor Hastie and Rob Tibshirani
| **[Book]** [Multivariate Data Visualization with R](http://lmdvr.r-forge.r-project.org/figures/figures.html) by Deepayan Sarkar
| **[Book]** [Machine Learning](https://bookdown.org/ssjackson300/Machine-Learning-Lecture-Notes/) by Samuel Jackson
| **[Book]** [Data Modeling Methods](https://bookdown.org/larget_jacob/data-modeling-methods/) by Jacob Larget
| **[Book]** [Introduction to R and Statistics](https://saestatsteaching.tech/) by University of Western Australia
| **[Book]** [Feature Engineering and Selection: A Practical Approach for Predictive Models](http://www.feat.engineering/index.html) by Max Kuhn and Kjell Johnson
| **[Book]** [Introduction to Research Methods](https://bookdown.org/ejvanholm/Textbook/) by Eric van Holm
| **[R Package]** [AppliedPredictiveModeling](https://cran.r-project.org/web//packages/AppliedPredictiveModeling/AppliedPredictiveModeling.pdf) by Max Kuhn
| **[R Package]** [caret](https://topepo.github.io/caret/index.html) by Max Kuhn
| **[R Package]** [rpart](https://mran.microsoft.com/web/packages/rpart/rpart.pdf) by Terry Therneau and Beth Atkinson
| **[R Package]** [lattice](https://cran.r-project.org/web/packages/lattice/lattice.pdf) by  Deepayan Sarkar
| **[R Package]** [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html/) by Hadley Wickham
| **[R Package]** [tidyr](https://cran.r-project.org/web/packages/tidyr/tidyr.pdf) by Hadley Wickham
| **[R Package]** [moments](https://cran.r-project.org/web/packages/moments/index.html) by Lukasz Komsta and Frederick
| **[R Package]** [skimr](https://cran.r-project.org/web/packages/skimr/skimr.pdf) by  Elin Waring
| **[R Package]** [RANN](https://cran.r-project.org/web/packages/RANN/RANN.pdf) by  Sunil Arya, David Mount, Samuel Kemp and Gregory Jefferis
| **[R Package]** [corrplot](https://cran.r-project.org/web/packages/corrplot/corrplot.pdf) by Taiyun Wei
| **[R Package]** [tidyverse](https://cran.r-project.org/web/packages/tidyverse/tidyverse.pdf) by Hadley Wickham
| **[R Package]** [lares](https://cran.rstudio.com/web/packages/lares/lares.pdf) by Bernardo Lares
| **[R Package]** [DMwR](https://mran.microsoft.com/snapshot/2016-05-02/web/packages/DMwR/DMwR.pdf) by Luis Torgo
| **[R Package]** [gridExtra](https://cran.r-project.org/web/packages/gridExtra/gridExtra.pdf) by Baptiste Auguie and Anton Antonov
| **[R Package]** [rattle](https://cran.r-project.org/web/packages/rattle/rattle.pdf) by Graham Williams
| **[R Package]** [RColorBrewer](https://cran.r-project.org/web//packages/RColorBrewer/RColorBrewer.pdf) by Erich Neuwirth
| **[R Package]** [stats](https://search.r-project.org/R/refmans/stats/html/00Index.html) by R Core Team

| **[R Package]** [caretEnsemble](https://cran.r-project.org/web/packages/caretEnsemble/caretEnsemble.pdf) by Zachary Deane-Mayer
| **[Article]** [A Brief Introduction to caretEnsemble)](https://cran.r-project.org/web/packages/caretEnsemble/vignettes/caretEnsemble-intro.html) by Zachary Deane-Mayer
| **[Article]** [How to Build an Ensemble Of Machine Learning Algorithms in R](https://machinelearningmastery.com/machine-learning-ensembles-with-r/) by Jason Brownlee
| **[Article]** [Ensemble Learning: Bagging, Boosting, and Stacking](https://towardsai.net/p/machine-learning/ensemble-learning-bagging-boosting-and-stacking) by Towards AI Team
| **[Article]** [Bagging, Boosting, and Stacking in Machine Learning](https://www.baeldung.com/cs/bagging-boosting-stacking-ml-ensemble-models) by Emmanuella Budu
| **[Article]** [Stacking Ensemble Machine Learning With Python](https://machinelearningmastery.com/stacking-ensemble-machine-learning-with-python/) by Jason Brownlee
| **[Article]** [Essence of Boosting Ensembles for Machine Learning](https://machinelearningmastery.com/essence-of-boosting-ensembles-for-machine-learning/) by Jason Brownlee
| **[Article]** [Ensemble Modeling with R](https://www.pluralsight.com/guides/ensemble-modeling-with-r) by Deepika Singh
| **[Article]** [Creating Ensemble Models in R](https://dustinrogers.github.io/Ensemble-Models/) by Dustin Rogers
| **[Publication]** [Ensemble Selection from Libraries of Models](https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.60.2859&rep=rep1&type=pdf) by Rich Caruana, Alexandru Niculescu-Mizil, Geoff Crew and Alex Ksikes (Proceedings of the 21 st International Conference on Machine Learning,)
| **[Course]** [Applied Data Mining and Statistical Learning](https://online.stat.psu.edu/stat508/) by Penn State Eberly College of Science
|
|
|
|